---
title: "Lab Assignment 04"
author: "Utkarsh Thusoo - 2020MCS120017"
date: "07/10/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(gapminder)
library(tidyverse)
library(ggplot2)
library(rmarkdown)
library(magrittr)
flying_etiquette <- read.csv("/media/sf_vmshare/iiitk/513/flying-etiquette.csv")
```



**Q1. Changing characters to factors?**

a. Change all character columns into factor columns.
```{r A1 (a)}
flying_etiquette %<>% mutate_if(is.character, as.factor)
str(flying_etiquette)
```

<br />

b. Remove people who responded “Never” to a question asking if they have flown before.

<br />

```{r A1 (b)}
missing <- which(flying_etiquette$How.often.do.you.travel.by.plane. == "Never", arr.ind = TRUE)
fe_1 <- flying_etiquette[-missing, ]
```

<br />

**Q2. Tidying data?**

a. Select columns where “rude” is in the column name.

```{r A2 (a)}
rude_columns <- which(str_detect(colnames(fe_1),"rude"))
rude_columns
```

<br />

b. Change the dataset from “wide” to “long”, with the variable names in a column called “response_var” and the values in a column called “value.”

```{r A2 (b)}
fe_2 <- gather(fe_1,response_var,value,rude_columns)
str(fe_2)
```

<br />

**Q3. Cleaning up strings?**

a. Use str_remove to remove everything before and including “rude to” (with the space at the end) in the response_var column.

```{r A3 (a)}
fe_3 <- fe_2 %>% mutate(response_var=str_remove(response_var,".*rude.to"))
str(fe_3)
```

<br />

b. Use str_remove to remove “on a plane” from the response_var column.

```{r A3 (b)}
fe_3 <- fe_3 %>% mutate(response_var=str_remove(response_var,'on.a.plane'))
str(fe_3)
```

<br />

**Q4. Dichotomizing variables?**

a. Remove rows with NA in the value column.

```{r A4 (a)}
fe_4 <- fe_3 %>% drop_na(value)
str(fe_4)
```

<br />

b. Create a new variable, rude, which is 0 if the value column is “No, not rude at all” or “No, not at all rude” and 1 otherwise.

```{r A4 (b)}
fe_4 <- fe_4 %>%  mutate(rude=ifelse(fe_4$value %in% c("No, not at all rude","No, not rude at all"),0,1))
str(fe_4)
```

<br />

**Q5. Summarizing data?**

a. Summarize the data set into two columns, the question (i.e. response_var), and a new column, perc_rude, the mean of the rude column for each question.  
b. Save it as rude_behaviors and then view your new dataset.

```{r A5 (a)}
rude_behaviors <-fe_4 %>% group_by(response_var) %>% summarise(perc_rude=mean(rude))
str(rude_behaviors)
```

<br />


**Q6. Create a plot?**

a. Using the dataset rude_behaviors in the previous question Order response_var by perc_rude.

```{r A6 (a)}
sorted_behaviour <- arrange(rude_behaviors, perc_rude)
```

b. Make a bar plot of response_var by perc_rude. Save it as initial_plot.

```{r A6 (b)}
initial_plot <- ggplot(sorted_behaviour, aes(x = reorder(response_var, perc_rude), y=perc_rude)) + geom_col() + labs(x = "response_var")
```

<br />

**Q7. Fixing Labels?**

a. Add its title, subtitle, and caption

```{r A7 (a)}
captioned_plot <- initial_plot + labs(title= "Hell is Other People in A Pressurized Metal Tube", subtitle = "Percentage of 874 air-passenger respondents who said action is very or somewhat rude", caption = "Source: Survey Monkey Audience")
captioned_plot
```

b. Set the labels for x and y-axis as empty strings.  
c. Save it as a new plot, titled_plot.

```{r A7 (b)}
titled_plot <- captioned_plot + labs(x = "", y = "")
titled_plot
```

<br />

**Q8. Flipping coordinates?**

a. Flip the x- and y-axis.

```{r A8 (a)}

flipped_plot <- titled_plot + coord_flip()
flipped_plot
```

b. Remove the x-axis ticks.

```{r A8 (b)}
flipped_plot + theme(axis.text.x = element_blank())

```

<br />

**Q9. Find out?**

a. The highest frequency of education

```{r A9 (a)}

tail(names(sort(table(flying_etiquette$Education))), 1)

```

b. The location where most people are from

```{r A9 (b)}
 tail(names(sort(table(flying_etiquette$Location..Census.Region.))), 1)
```

<br />
